### 2025-05-29 docs/article/Sections/conclusion.tex
article 50%~ completed
Completed Article disclaimers

\documentclass[../apG48.tex]{subfiles}

\begin{document}

Through benchmarking across Spring WebFlux, Spring MVC, and Quarkus, we evaluated eight different template engines and found that non-blocking implementations—especially those using virtual threads—consistently deliver performance on par with traditional blocking approaches under high concurrency. 
These results highlight virtual threads as a promising alternative to complex asynchronous programming models, offering a simpler development experience without sacrificing scalability or responsiveness.

The next steps include a deeper analysis of the benchmark results in the final report, as well as expanded coverage of the current state of non-blocking PSSR and asynchronous programming techniques. Additionally, the project may be extended to support other frameworks such as Vert.x.

\end{document}

>>>>

\section{Conclusion}

In recent decades, non-blocking I/O has become the standard approach for building
highly responsive and scalable web servers. However, traditional synchronous
programming models are not compatible with non-blocking APIs, which typically
rely on callback-based conventions such as
\textit{continuation-passing style} (CPS)~\cite{scheme} or
\textit{promises}~\cite{promise}. These approaches not only hinder
sequential readability but also increase code verbosity, making them more
error-prone.
Alternatives like the \texttt{async}/\texttt{await} idiom~\cite{async_await}
or \textit{suspend functions}~\cite{elizarov2021coroutines} simplify
asynchronous programming by mimicking a sequential style without blocking
threads.
Recent proposals~\cite{carvalho2023async,wise2024pssr} have applied contemporary
asynchronous idioms to SSR web templates, demonstrating how they can overcome
the scalability bottlenecks present in traditional web template engines.

As an alternative, Java Virtual Threads can be applied to any blocking I/O call,
leveraging the Java runtime to transparently intercept and convert it into non-blocking I/O,
without requiring any changes to the calling code or exposing its internal complexities.
In this work, we explored how this technique can be applied to traditional
web template engines and whether it can achieve performance competitive with
reactive approaches provided by frameworks such as Thymeleaf~\cite{webflux} and HtmlFlow~\cite{htmlflow}.
Our benchmarks across Spring WebFlux, Spring MVC, and Quarkus show that synchronous
non-blocking execution using virtual threads consistently delivers performance
comparable to asynchronous non-blocking approaches under high concurrency.
These findings highlight virtual threads as a promising alternative to complex
asynchronous programming models, offering a simpler development experience
without compromising scalability or responsiveness.

### 2025-05-29 docs/article/Sections/introduction.tex
Completed Article disclaimers

\textit{Low-thread} servers, also known as
\textit{event-driven}~\cite{event-driven-servers}, have gained prominence in
contemporary web applications due to their ability to efficiently manage a large
number of concurrent I/O operations with minimal resources, thus promoting
better scalability.
By leveraging asynchronous I/O operations—such as
database queries and API calls, servers can avoid blocking threads while
waiting for data, thereby maximizing throughput and responsiveness. To support
this non-blocking architecture, PSSR implementations require template engines
that are compatible with asynchronous data models~\cite{carvalho2023async}.
Some modern template
engines, such as HtmlFlow~\cite{htmlflow} and Thymeleaf~\cite{thymeleaf}, have
been designed with these capabilities in mind. However, many traditional
template engines—particularly those using external domain-specific languages
(DSLs)~\cite{Fowler03}—still depend on blocking interfaces like
\texttt{Iterable} for data processing. This blocking behavior forces server
threads to remain idle until the entire HTML output is ready, undermining the
performance benefits of non-blocking I/O and limiting scalability in
high-concurrency environments.

With the introduction of \textit{Virtual Threads}~\footnote{\url{https://openjdk.org/jeps/444}}
in Java 21, it is now possible to execute blocking I/O operations in a
scalable, lightweight manner. This capability allows traditional template
engines—often reliant on blocking interfaces—to operate efficiently in
high-concurrency, non-blocking environments without requiring complex
asynchronous programming models. As a result, PSSR
can now be implemented using familiar HTML templates, 
simplifying development and improving maintainability.

>>>>

\textit{Low-thread} servers, also known as
\textit{event-driven}~\cite{event-driven-servers}, have gained prominence in
contemporary web applications due to their ability to efficiently manage a large
number of concurrent I/O operations with minimal resources, thus promoting
better scalability.
By leveraging asynchronous I/O operations—such as
database queries and API calls, servers can avoid blocking threads while
waiting for data, thereby maximizing throughput and responsiveness. To support
this non-blocking architecture, PSSR implementations require template engines
that are compatible with asynchronous data models~\cite{carvalho2023async}.
Some modern template
engines, such as HtmlFlow~\cite{htmlflow} and Thymeleaf~\cite{thymeleaf}, have
been designed with these capabilities in mind. However, many legacy
template engines—particularly those using external domain-specific languages
(DSLs)~\cite{Fowler03}—still depend on blocking interfaces like
\texttt{Iterable} for data processing. This blocking behavior forces server
threads to remain idle until the entire HTML output is ready, undermining the
performance benefits of non-blocking I/O and limiting scalability in
high-concurrency environments.

With the introduction of \textit{Virtual Threads}~\footnote{\url{https://openjdk.org/jeps/444}}
in Java 21, it is now possible to execute blocking I/O operations in a
scalable, lightweight manner. This capability allows legacy template
engines—often reliant on blocking interfaces—to operate efficiently in
high-concurrency, non-blocking environments without requiring complex
asynchronous programming models. As a result, PSSR
can now be implemented using familiar HTML templates, 
simplifying development and improving maintainability.
